---
title: "Business Intelligence Project"
author: "<Specify your name here>"
date: "<16/10/2023>"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

|                                              |     |
|----------------------------------------------|-----|
| **Student ID Number**                        |135791,136657,136665 |
| **Student Name**                             | Rehema Jemutai , Shirleen Murugi, Adnan Mwendwa |
| **BBIT 4.2 Group**                           | C,B |
| **BI Project Group Name/ID (if applicable)** | ... |

# Setup Chunk

**Note:** the following KnitR options have been set as the global defaults: <BR> `knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy = TRUE)`.

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, include=FALSE}
library(formatR)
knitr::opts_chunk$set(
  warning = FALSE,
  collapse = FALSE
)
```

# Understanding the Dataset (Exploratory Data Analysis (EDA))

## Loading the Dataset

### Source:

The dataset that was used can be downloaded here: *\<provide a link\>*

### Reference:

*\<Cite the dataset here using APA\>\
Refer to the APA 7th edition manual for rules on how to cite datasets: <https://apastyle.apa.org/style-grammar-guidelines/references/examples/data-set-references>*

```{r Dataset Loader}
library(readr)

if (require("languageserver")) {
  require("languageserver")
} else {
  install.packages("languageserver", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

if (require("mlbench")) {
  require("mlbench")
} else {
  install.packages("mlbench", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## readr ----
if (require("readr")) {
  require("readr")
} else {
  install.packages("readr", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## caret ----
if (require("caret")) {
  require("caret")
} else {
  install.packages("caret", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## e1071 ----
if (require("e1071")) {
  require("e1071")
} else {
  install.packages("e1071", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## factoextra ----
if (require("factoextra")) {
  require("factoextra")
} else {
  install.packages("factoextra", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## FactoMineR ----
if (require("FactoMineR")) {
  require("FactoMineR")
} else {
  install.packages("FactoMineR", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}



```


```{r Step 2}
library(readr)
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset <- read_csv("../data/20230412-20230719-BI1-BBIT4-1-StudentPerformanceDataset.csv")
View(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)

```

```{r step 3}
library(caret)

summary(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_spiritual_wellness <- as.numeric(unlist(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset[, 41]))
hist(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_spiritual_wellness, main = names(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)[41])

model_of_the_transform <- preProcess(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset, method = c("scale"))
print(model_of_the_transform)
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_scale_transform <- predict(model_of_the_transform, X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)

summary(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_scale_transform)
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_spiritual_wellness <- as.numeric(unlist(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_scale_transform[, 41]))
hist(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_spiritual_wellness, main = names(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_scale_transform)[41])


```


```{r Step 4}
summary(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)
model_of_the_transform <- preProcess(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset, method = c("center"))
print(model_of_the_transform)
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_center_transform <- predict(model_of_the_transform, X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)
summary(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_center_transform)
```


```{r Step 5}
library(readr)

summary(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset[, 4], sd)
model_of_the_transform <- preProcess(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset,
                                     method = c("scale", "center"))
print(model_of_the_transform)
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_standardize_transform <- predict(model_of_the_transform, X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset) # nolint

# AFTER
summary(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_standardize_transform)
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_standardize_transform[, 4], sd)
```

```{r step 6}
library(readr)

summary(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)
model_of_the_transform <- preProcess(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset, method = c("range"))
print(model_of_the_transform)
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_normalize_transform <- predict(model_of_the_transform, X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)
summary(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_normalize_transform)
```

```{r step 7}

summary(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_standardize_transform)

# Calculate the skewness before the Box-Cox transform
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_standardize_transform[, 4],  skewness, type = 2)
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_standardize_transform[, 4], sd)

model_of_the_transform <- preProcess(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_standardize_transform,
                                     method = c("BoxCox"))
print(model_of_the_transform)
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_box_cox_transform <- predict(model_of_the_transform,
                                                                                      X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_standardize_transform)

# AFTER
summary(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_box_cox_transform)

sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_box_cox_transform[, 4],  skewness, type = 2)
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_box_cox_transform[, 4], sd)

# Calculate the skewness after the Box-Cox transform
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_box_cox_transform[, 4],  skewness, type = 2)
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_box_cox_transform[, 4], sd)
```


```{r step 8}
summary(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_standardize_transform)

sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_standardize_transform[, 4],  skewness, type = 2)
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_standardize_transform[, 4], sd)

model_of_the_transform <- preProcess(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_standardize_transform,
                                     method = c("YeoJohnson"))
print(model_of_the_transform)
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_yeo_johnson_transform <- predict(model_of_the_transform, # nolint
                                           X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_standardize_transform)

# AFTER
summary(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_yeo_johnson_transform)


sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_yeo_johnson_transform[, 4],  skewness, type = 2)
sapply(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_yeo_johnson_transform[, 4], sd)
```






```{r step 9a}


```

```{r step 9b}
```



```{r step 10}

if (!is.element("fastICA", installed.packages()[, 1])) {
  install.packages("fastICA", dependencies = TRUE)
}
require("fastICA")


summary(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)
model_of_the_transform <- preProcess(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset,
                                     method = c("scale", "center", "ica"),
                                     n.comp = 3)
print(model_of_the_transform)
X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_ica_dr <- predict(model_of_the_transform, X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset)

summary(X20230412_20230719_BI1_BBIT4_1_StudentPerformanceDataset_ica_dr)


```
